Molecular Gas Kinematics of the Strongly-lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-11-18 WRITEUP/ms.tex
only keep necessary stuff for recoil

\subsection{Systemic Redshift and Velocity Offset}
%Outflow:
% Not AGN outflows because require different velocity offsets in Balmer and forbidden lines (Zamanov02a)
% line origins:
% \mgii & Hbeta: outer part of the BLR (Goad12a) compared to high-ionization lines (\eg CIV)
% Hbeta can also come from NLR

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim0.654, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and
the \mgii emission line.
Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent with the systemic redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.

% REFEREE: offset in BLR and NLR in RXJ1131 has been known before, not new for this paper
%One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
%the central AGN, whereas this condition is not necessary for NLRs.
%From an SDSS sample of $>$3800 quasars, \citet{Richards02a} find a
%median of 97\kms and a dispersion of 269\,\kms in the distribution of
%the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines, suggesting
%that large offsets are uncommon. Nevertheless,
%offsets as large as $\sim$\,500\,\kms have been observed in some cases.
%%% In an SDSS sample of ongoing mergers at $z$\,$<$\,0.21, AGN lines are offset by $\lesssim$410\,km/s
%%\citep[\eg][]{Richards02a, Comerford14a}. % between AGN lines (forbidden, Balmer) and stellar absorption

%Alternatively, large velocity offsets between emission lines of the
%BLR and the NLR may arise in a recoiling black hole (BH),
%% recoiling SMBHs as well as subparsec-scale binary SMBHs produce velocity offsets in the broad AGN emission lines and not the narrow AGN emission lines (e.g., Gaskell 1983; Bonning et al. 2007; Eracleous et al. 2012; Ju et al. 2013; Shen et al. 2013).
%where its BLR is moving at high velocity relative to the bulk of its host galaxy
%\citep{Madau04a, Bonning07a, Loeb07a}.
%A recoiling black hole is expected
%when a pair of uneven mass BH coalesce, during which their orbital energy is being released
%as gravitational wave and a
%non-zero net angular momentum is being carried away.
%Depending on their initial conditions, numerical relativity simulations have shown that
%the recoil velocity can reach up to $\sim$4000\,\kms for spinning BHs \citep[\eg][]{Campanelli07a}.
%Several sources have been proposed as recoiling BH candidates, such as CID-42 at $z$\,=\,0.359 \citep{Civano10a} and
%SDSS 0956+5127 at $z$\,$\sim$\,0.714 \citep{Steinhardt12a}.
%Since this model requires a merged BH, it would imply that RXJ1131 might
%have been in a previous merger, which would also explain
%the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

>>>>

\subsection{Systemic Redshift and Velocity Offset}
%Outflow:
% require different velocity offsets in Balmer and forbidden lines (Zamanov02a)
% line origins:
% \mgii & Hbeta: outer part of the BLR (Goad12a) compared to high-ionization lines (\eg CIV)
% Hbeta can also come from NLR

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim0.654, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and the \mgii emission line.
Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent with the systemic redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.
Offset between BLR and NLR lines have been reported in literature, e.g., 
\citet{Richards02a} find a
median of 97\kms and a dispersion of 269\,\kms in the distribution of
the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in an SDSS sample of $>$3800 quasars, 
and \citet{Bonning07a} find a mean offset of 100\,\kms with a dispersion of 212\,\kms %FWHM = 500
between the broad component of
H$\beta$ and \oiii lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.81,
illustrating that large offsets ($\Delta v$$\gtrsim$\,500\,\kms) are uncommon but have been observed in some cases.

Common interpretations invoked to explain offsets between BLR and systemic redshifts include
e.g., outflow/inflow of gas in the BLR, rotation of the accretion disk, viewing angle towards the accretion disk, 
inclination effect, in which clumps in the accretion disk may block emission towards the far side of the disk, and thus the observed BLR emission is only
originating from
the redshifted inflowing gas component in the near side of the accretion disk of RXJ1131.
In the latter, \citet{Sluse07a} report a covering factor of 20\% based on \mgii absorption line at $z$\eq0.654.
Additionally, differentially lensing of the compact BLR emission by microlensing in RXJ1131\citep[\eg][]{Sluse07a, Sluse12a} 
may have boosted the redshifted component more than the blueshifted component,
thereby biasing the BLR line centroids towards longer wavelengths.
{\bf DR: is it because of the large linewidth (one component of \mgii is $>$20000\kms, and broad comp. of H$\beta$ is $\sim$2000$-$4000\kms) that the referee thinks it's more likely to be inflow and therefore doesn't like the recoil black hole?
Sluse07:
"Another remarkable result is the presence in our spectra of a \mgii absorption doublet blueshifted at $z$\eq0.654. 
According to its velocity shift ($\Delta v$\ssim$-$660\,\kms). 
These lines disappear from the spectrum ratios A/B and C/B. This indicates that the absorbed flux is proportional to the flux coming from the continuum+BLR. This implies that the region at the origin of the absorption must cover both the continuum and the BLRs and that, within the uncertainties, their depths are identical in the spectra of the three images A, B, and C."
ok.. so is it because of this broad absorption component of \mgii that is at the same $z$ as the NLR lines, and therefore implying that the BLR and 
the AGN is not really spatially and kinematically offset, 
but the offset is truly just a manifestation of a rotating disk + other kinematics (+ absorption of continuum by BLR moving towards us, redshfiting the line centroid, + inflow of material + clumps blocking the far side inflow?)
}

Alternatively, the observed offset %between the BLR and the systemic redshift (and the NLR)
may be explained by a more extreme scenario of a recoiling black hole (BH), 
where its BLR is moving at high velocity relative to the bulk of its host galaxy
\citep{Madau04a, Bonning07a, Loeb07a}.
% where
% BLRs are gravitationally bound to
% the AGN (with a BLR dynamical timescale $\sim$10\E{2} years \citep{Bonning07a}, \ie can be quickly regenerated), 
% whereas this condition is not necessary for NLRs.
%
% In hierarchical model of galaxy formation, supermassive black holes are expected to form through
% mergers, of which the majority took place at high redshifts. 
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a non-zero net angular momentum is being carried away.
Depending on their initial conditions (e.g., black hole mass ratio, spin-orbit orientation, spin magnitude), 
numerical relativity simulations have shown that the recoil velocity can reach up to 
$v_{\rm kick}$\ssim4000\,\kms for spinning BHs, % \citep[\eg][]{Campanelli07a}, 
with a typical offset recoil velocity of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\citep[\eg][]{Libeskind06a,Campanelli07a}.
% ------------- fate ----------------
% Several outcomes of the recoiling blackholes include,
% a) In spherical bulges the black hole will oscillate right through the center, and return to the center after a few dynamical times, or 
% deplete or "heat up" the stars in the nucleus, reducing the efficiency of dynamical and extending the return time.
% b) If the black hole is able to leave the bulge / ejected and enters a triaxial dark matter halo, it may not return at all within a Hubble time \citep{Guedes09a} and become intergalactic/naked black holes (but rare unless in recoil velocity $>>$ few thousands \kms, i.e., sufficient to escape the host galactic potential) \citep[\eg][]{Loeb07a}.
%c) High resolution simulations of gas mergers (e.g. Mayer et al. 2007) show that up to 60\% of the initial gas is funneled to the center of the remnant. In that particular simulation, the mass of the nuclear disk is Md = 3e9 Msun, and recoliing black holes with velocities of 
%$~$500 km/s are not able to escape the central region (scale length L=75 pc).
% In the case of gas-rich merger remnants, where large amounts of gas have been
% funneled to the central regions find that the MBHs remain within R$<$1\,kpc from the center
% of the host even for high recoil velocities ($v_{\rm kick}$\eq1200\,\kms) due to the compactness of the
% remnant galaxy's nuclear disk" \citep{Mayer07a,Guedes08a}.
% ----------------------------------------
%A recoiling black hole may therefore 
%oscillate around the center of the galaxy as its orbit is decaying due to dynamical friction, 
%or leave its host galaxy if $v_{\rm kick}$$\gg$$v_{\rm esc}$ and dynamical friction is inefficient 
%\citep[see \eg \citealt{Volonteri10a} for a discussion]{Guedes09a,Bonning07a, Colpi14a}. 
%% escape velocity
%If we assume a \citet{Hernquist90a} profile for the stellar bulge in RXJ1131, 
%%$M_{\rm BH}$/\mstar\ssim1/375, $v_{\rm circ}$\ssim$v_{\rm rot}, 
%% then the upper limit of $v_{\rm kick}$\ssim2$\times$$v_{\rm circ}$, 
%% bulge escape velocity $v_{\rm esc}$\eq2.83$\v_{\rm circ}$ = 858\kms
%% %(in the case of a density profile of the spherical \citet{Hernquist90a} model for the stellar bulge, 
%% %spherically symmetric isotropic model)%  \citep{BoylanKolchin04a}, 
%then the escape velocity $v_{\rm esc}$\eq2$\sqrt{2}$$v_{\rm circ}$, where $v_{\rm circ}$ is the velocity
%at $r$\eq$a$, the stellar bulge scale length that is at most a few kpc.
%Assuming that the rotation curve of RXJ1131 is similar to that MW, then at inner radius ($\lesssim$ few kpc), 
%we would expect $v_{\rm rot} (r=a)\gtrsim$$v_{\rm rot}$ of the extended gas disk (see \Sec{dynamics}).
%Given that we don't know precisely the stellar bulge scale length, if we assume 
%$v_{\rm rot} (r=a)$\ssim$v_{\rm rot}$ of the gas disk, then $v_{\rm esc}$\ssim860\,\kms. 
%Therefore, the recoil BH in RXJ1131 is unlikely to escape its gas-rich host galaxy.%  \citep[\eg][]{Loeb07a, Volonteri10a}.
Several sources have been proposed as recoiling BH candidates, such as CID-42 at $z$\,=\,0.359 \citep{Civano10a},
SDSS 0956+5127 at $z$\,$\sim$\,0.714 \citep{Steinhardt12a}, and
SDSSJ0927+2943 at $z$\,$\sim$\,0.713, where a set of redshifted narrow lines and a second set of blueshifted broad AGN lines
have been observed \citep{Komossa08a}.
However, this scenario for SDSSJ0927+2943 was recently refuted by \citet{Decarli14a}, who find that the
BLR lines in SDSSJ0927+2943 are indeed consistent with the CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm the redshifted BLR.
Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of previous merger, which is not implausible and 
might also explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

The available data do not provide direct evidence favoring any of the above scenarios, 
we therefore cannot rule out any of the above plausible scenarios; $\lesssim$1-kpc
resolution observations probing the morphology, structures and kinematics of the (ionized) gas and 
stellar (e.g., via H$\alpha$ kinematics measurements) would be useful to distinguish between these scenarios.
